Hierarchy Builder: algebraic~hierarchies made easy in Coq with Elpi

### 2020-04-27 hb.tex
Explain the flat variant of packed classes

Next we define Abelian groups.
Since the monoid structure is the bottom of this hierarchy its class record
\coq{Monoid} consists just one \mixin{}. In contrast the class record
of Abelian groups consists of two \mixins{} where the second one depends
 on the former (since Abelian groups inherit from monoids).
\begin{coqcode}
Module AbelianGroup.
Record axioms (T : Type) : Type := Class {
  Monoid_of_Type_mixin : Monoid_of_Type T;
  AbelianGroup_of_Monoid_mixin : AbelianGroup_of_Monoid T Monoid_of_Type_mixin; }.
Structure type : Type := Pack { sort : Type; class : axioms sort }.
End AbelianGroup.
\end{coqcode}
As in the monoid structure, we declare \coq{AbelianGroup.sort} as an implicit
coercion and lift additive inverse \coq{opp} and Abelian group axioms.
Since Abelian groups inherit from monoids, we also declare an implicit coercion
from Abelian groups to monoids. A coercion between structures can be defined
in two steps: first a coercion between the class record of the
superclass to the class record of the subclass (line~\ref{demo4:coe:ag:monoid:class});
and a coercion between
structure records (line~\ref{demo4:coe:ag:monoid:type}) relying on the first one.
\begin{coqcode}
Coercion AbelianGroup.sort : AbelianGroup.type >-> Sortclass.
Coercion AbelianGroup_class_to_Monoid_class (T : Type) (c : AbelianGroup T) : #\label{demo4:coe:ag:monoid:class}#
  Monoid T := Monoid.Class T (AbelianGroup.Monoid_of_Type_mixin T c).
Coercion AbelianGroup_to_Monoid (T : AbelianGroup.type) : Monoid.type := #\label{demo4:coe:ag:monoid:type}#
  Monoid.Pack (AbelianGroup.sort T) (AbelianGroup.class T).

@@

\subsection{Multiple inheritance}
In order to introduce multiple inheritance, we extend the hierarchy described in
Section~\ref{subsec:v3} with the structure of semirings %, which is another
%intermediate structure between monoids and rings
as depicted in Figure~\ref{fig:puzzle_v4}, and name it V4. Semirings introduce the
binary multiplication operator \coq{mul} and its identity element
\coq{one}.  For the sake of completeness, the full code of V4 is available in
Appendix~\ref{sec:full-v4-code}.

\begin{coqcode}
#\hbmixin{}# Record SemiRing_of_Monoid A of Monoid A := {
  one : A;
  mul : A -> A -> A;
  (* 7 axioms are omitted. *)
}.
\end{coqcode}

\begin{wrapfigure}[3]{r}{.4\textwidth}
  \includegraphics[width=.4\textwidth]{v4.pdf}
  \caption{V4 introduces multiple inheritance. For brevity we omit the factories/builders for the upper arrows.}
  \label{fig:puzzle_v4}
\end{wrapfigure}
\
\begin{coqcode}
#\hbfactory{}# Record Ring_of_AbelianGroup A
  of AbelianGroup A := {
  (* 2 operators and 5 axioms are omitted. *)
}.
#\hbbuilders{}# Context A (a : Ring_of_AbelianGroup A).
..
#\hbend{}#.
\end{coqcode}

>>>>

Next we define Abelian groups.
Since the monoid structure is the bottom of this hierarchy its class record
\coq{Monoid} consists of just one \mixin{}. In contrast the class record
of Abelian groups consists of two \mixins{} where the second one depends
 on the former (since Abelian groups inherit from monoids).
\begin{coqcode}
Module AbelianGroup.
Record axioms (T : Type) : Type := Class {
  Monoid_of_Type_mixin : Monoid_of_Type T;
  AbelianGroup_of_Monoid_mixin : AbelianGroup_of_Monoid T Monoid_of_Type_mixin; }.
Structure type : Type := Pack { sort : Type; class : axioms sort }.
End AbelianGroup.
\end{coqcode}
In the flat variant of Packed Classes, a class record gathers mixins as its
fields directly as in above. In contrast, a class record of the non-flat variant
packs a class instance of one of the superclasses with remaining mixins, which
reduces the amount of code to implement inheritance significantly.
Nested class records and chains of their projections exponentially slow down the
conversion while the flat variant can mitigate the issue. Since we do not need
to care about the amount of code in automated generation, \hb{} uses the flat
variant of Packed Classes as its target language.

As in the monoid structure, we declare \coq{AbelianGroup.sort} as an implicit
coercion and lift additive inverse \coq{opp} and Abelian group axioms.
Since Abelian groups inherit from monoids, we also declare an implicit coercion
from Abelian groups to monoids. A coercion between structures can be defined
in two steps: first a coercion between the class record of the
superclass to the class record of the subclass (line~\ref{demo4:coe:ag:monoid:class});
and a coercion between
structure records (line~\ref{demo4:coe:ag:monoid:type}) relying on the first one.
\begin{coqcode}
Coercion AbelianGroup.sort : AbelianGroup.type >-> Sortclass.
Coercion AbelianGroup_class_to_Monoid_class (T : Type) (c : AbelianGroup T) : #\label{demo4:coe:ag:monoid:class}#
  Monoid T := Monoid.Class T (AbelianGroup.Monoid_of_Type_mixin T c).
Coercion AbelianGroup_to_Monoid (T : AbelianGroup.type) : Monoid.type := #\label{demo4:coe:ag:monoid:type}#
  Monoid.Pack (AbelianGroup.sort T) (AbelianGroup.class T).

@@

\subsection{Multiple inheritance}

\begin{wrapfigure}[6]{r}{.4\textwidth}
 \includegraphics[width=.4\textwidth]{v4.pdf}
 \caption{V4 introduces multiple inheritance. For brevity we omit the factories/builders for the upper arrows.}
 \label{fig:puzzle_v4}
\end{wrapfigure}
In order to introduce multiple inheritance, we extend the hierarchy described in
Section~\ref{subsec:v3} with the structure of semirings %, which is another
%intermediate structure between monoids and rings
as depicted in Figure~\ref{fig:puzzle_v4}, and name it V4. Semirings introduce the
binary multiplication operator \coq{mul} and its identity element
\coq{one}.  For the sake of completeness, the full code of V4 is available in
Appendix~\ref{sec:full-v4-code}.

\begin{coqcode}
#\hbmixin{}# Record SemiRing_of_Monoid A
  of Monoid A := {
  one : A;
  mul : A -> A -> A;
  (* 7 axioms are omitted. *)
}.

#\hbfactory{}# Record Ring_of_AbelianGroup A
  of AbelianGroup A := {
  (* 2 operators and 5 axioms are omitted. *)
}.
#\hbbuilders{}# Context A (a : Ring_of_AbelianGroup A).
..
#\hbend{}#.
\end{coqcode}
